---
title: "reproducible research - PA1"
author: "lkomenda"
date: "Sunday, December 14, 2014"
output: html_document
---

###Loading and preprocessing the data


```{r}
read.csv("activity.csv", as.is = 2)->data1
```
some summary of the datafile
```{r}
str(data1)
```


###What is mean total number of steps taken per day?
prepare data
```{r}
steps_per_day<-aggregate(data1$steps, by = list(data1$date), 
          FUN = function(x) sum(x, na.rm = T))
colnames(steps_per_day)<-c("day", "sum_of_steps")
```

histogram of total number of steps per day:
```{r}
hist(steps_per_day$sum_of_steps, breaks = 10)
```

median and arithmetic mean are:
```{r}
summary(steps_per_day$sum_of_steps)[3:4]
```

###What is the average daily activity pattern?
prepare data
```{r}
steps_per_interval<-aggregate(data1$steps, by = list(data1$interval), 
          FUN = function(x) mean(x, na.rm = T))
colnames(steps_per_interval)<-c("interval", "mean_of_steps")
```

histogram of total number of steps per day:
```{r}
plot(steps_per_interval$interval, steps_per_interval$mean_of_steps, type ="l",
main = "plot of interval and average steps taken",
xlab = "interval", ylab = "mean steps")
```

Which interval has maximum steps
```{r comuputing max}
max1<-max(steps_per_interval$sum_of_steps)
w1<-which(steps_per_interval$sum_of_steps==max1)
interval_max<-steps_per_interval$interval[w1]
```

maximum is `r max1` and it happens at interval `r interval_max`.

###Imputing missing values

computing uncomplete cases
```{r}
ok_case<-complete.cases(data1)
not_ok_case_n<-sum(!ok_case)
```
The number of uncoplete is `r not_ok_case_n`

Prepare the dataset with medians for each interval
```{r}
steps_median<-aggregate(data1$steps, by = list(data1$interval), 
          FUN = function(x) median (x, na.rm = T))
colnames(steps_median)<-c("interval", "median_of_steps")
```

merge medians to old data and repleace missings with it

```{r}
merge (data1, steps_median, by  = "interval", all.x = T)->data2
data2[!ok_case, "steps"]<-data2[!ok_case, "median_of_steps"]
data2<-data2[,1:3]
```

prepare data
```{r}
steps_per_day2<-aggregate(data2$steps, by = list(data2$date), 
          FUN = function(x) sum(x, na.rm = T))
colnames(steps_per_day2)<-c("day", "sum_of_steps")
```

histogram of total number of steps per day:
```{r}
hist(steps_per_day2$sum_of_steps, breaks = 10)
```

median and arithmetic mean are:
```{r}
summary(steps_per_day2$sum_of_steps)[3:4]
```

the differences between raw and imputed data summaries is:

```{r}
summary(steps_per_day$sum_of_steps)[3:4] - 
      summary(steps_per_day2$sum_of_steps)[3:4]
```
so both mean and median are lower after median substitution

###Are there differences in activity patterns between weekdays and weekends?

first prepare data, we will set locale to us to have english weekday names,
I hope it will work on non Windows machines.
```{r}
Sys.setlocale("LC_TIME", "US")
data1$weekday<-weekdays(as.Date(data1[,2]))
weekend_days<-c("Saturday", "Sunday")
data1$weekday[data1$weekday %in% weekend_days]<-"weekend"
data1$weekday[data1$weekday!= "weekend"]<-"weekday"
data1$weekday<-as.factor(data1$weekday)

```

let's prepare aggregate data for ploting
```{r}
steps_per_interval_weekday<-aggregate(data1$steps, by = list(data1$weekday, data1$interval), 
          FUN = function(x) mean(x, na.rm = T))
colnames(steps_per_interval_weekday)<-c("weekday", "interval", "mean_of_steps")
```


plot steps timeseries for weekdays and weekends

```{r, fig.height= 6}
par(mfrow=c(2,1))
plot(steps_per_interval_weekday$interval[steps_per_interval_weekday$weekday=="weekday"], 
     steps_per_interval_weekday$mean_of_steps[steps_per_interval_weekday$weekday=="weekday"], type ="l",
main = "plot of interval and average steps taken in weekdays",
xlab = "interval", ylab = "mean steps")

plot(steps_per_interval_weekday$interval[steps_per_interval_weekday$weekday!="weekday"], 
     steps_per_interval_weekday$mean_of_steps[steps_per_interval_weekday$weekday!="weekday"], type ="l",
main = "plot of interval and average steps taken in weekends",
xlab = "interval", ylab = "mean steps")

```
